Dynastic Extinctions and Urban Economic Development: \\Evidence from the Holy Roman Empire\footnote{I thank Davide Cantoni, Matthias Weigand, and my RA colleagues at the Chair of Economic History for helpful comments and feedback.}

### 2023-06-10 paper/paper.tex

My story is not about population but institutions (although it is compatible with Malthusianism). A longstanding tradition in economics emphasises the role of institutional factors in long-run growth (e.g. \cite{north1970}, \cite{delong1993}, \cite{ajr2001}). From this vast literature one theme is especially relevant here: the institution of fiscal capacity, or the state's ability to collect taxes. \cite{tilly1985} argues that the main reason for states to collect taxes is to be able to wage war. In other words, fiscal capacity and state power are tightly linked --- so if conflict selects for state power, then it also selects for fiscal capacity. \cite{cantoni2023} show that conflict drove state consolidation. \cite{dincecco2012} show that fiscal capacity had a positive effect on growth.

@@

Empirical tests have suffered from a lack of available historical data. Such a study requires at the very least three things: 1) a detailed territorial history of European cities, 2) some city-level measure of development reaching far enough into the past, and 3) some measure of state power, institutional quality, or whatever else one wishes to explain effects with. These data limitations are starting to be overcome.

>>>>

My story is not about population but institutions (although it is compatible with Malthusianism). A longstanding tradition in economics emphasises the role of institutional factors in long-run growth (e.g. \cite{north1970}, \cite{delong1993}, \cite{ajr2001}). From this vast literature one theme is especially relevant here: the institution of fiscal capacity, or the state's ability to collect taxes. \cite{tilly1985} argues that the main reason for states to collect taxes is to be able to wage war. In other words, fiscal capacity and state power are tightly linked --- so if conflict selects for state power, then it also selects for fiscal capacity. \cite{gennaioli2015} and \cite{cantoni2023} show that conflict drove state consolidation. \cite{dincecco2012} show that fiscal capacity had a positive effect on growth.

@@

I use data from the \textit{Princes and Townspeople} database which is largely based on the \textit{Deutsches Städtebuch} \citep{keyser1939} and \cite{kobler2007}'s \textit{Historisches Lexikon der Deutschen Länder}. The specific datasets I use are those on city locations \citep{pt1}, territorial histories \citep{pt2}, construction activity \citep{pt5} and conflict incidents \citep{pt6}.\footnote{I thank Davide Cantoni and Matthias Weigand for giving me access to unpublished yearly versions of the construction and territorial histories data.} I aggregate data from these sources to obtain a city-by-year panel of switching events, construction activity, and conflict.

Switching: describe format; total number of recorded switches; territory id.

Construction: describe format / split and rationale. This can be two paragraphs.

Conflict: describe format (just a dummy).

I also build coarsened versions of this dataset in which I aggregate years into longer periods. Apart from the yearly version there is also a 10-yearly version, a 50-yearly version and a 100-yearly version. Not only does this make event studies nicer to look at, it also allows me to use more of the conflict and construction data: the 50-year and 100-year versions contain events that aren't recorded with a precise year, but only "beginning of the 16th century" and the like.
